Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-06-18 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: responded to comments #1 and #3 in referee report.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{latticepts.eps}
    \caption{A two-dimensional projection of lattice points for increasing sample size.  The later points fill in gaps because they are highly correlated with the earlier points.  These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}. }
    \label{CDHJSfiglatticepts}
\end{figure}

@@

Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.  Stephen Joe and Frances Kuo \cite{JoeKuo03,KuoJoe08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.  Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria. During lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these implementations in large libraries would benefit a wider menu of LD offerings, including custom generators from LatNetBuilder or other sources.  Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings. SAS facilitates flexible integration of third-party open-source software; see \href{https://www.sas.com/en_us/software/viya/open.html}{\url{https://www.sas.com/en_us/software/viya/open.html}}.
This presents opportunities for developers of smaller libraries with more QMC features, like BRODA and QMCPy, to connect their libraries with larger software environments. An example of QMCPy connecting with other software is given in the next subsection.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five. These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{CDHJSfigbeam}
\end{figure}

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.  We illustrate in Figure \ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.  The beam position is discretized at $31$ points.  QMCPy computes the expected deflection of the beam at these points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

>>>>

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{latticepts.eps}
    \caption{A two-dimensional projection of lattice points for increasing sample size.  The later points fill in gaps because they are highly correlated with the earlier points.  These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}. }
    \label{CDHJSfiglatticepts}
\end{figure}

@@

Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.  Stephen Joe and Frances Kuo \cite{JoeKuo03,KuoJoe08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.  Pierre L'Ecuyer and collaborators have created LatNet Builder \cite{LEcEtal22a} to construct good lattice and polynomial lattice sequences based on user-defined criteria. During lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries. \SCComment{Changed LatNet Builder's reference in response to Referee's \#3, but the citation relies on another reference ArnGilKel22a which is not found}

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these implementations in large libraries would benefit a wider menu of LD offerings, including custom generators from LatNet Builder or other sources.  Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings. SAS facilitates flexible integration of third-party open-source software; see \href{https://www.sas.com/en_us/software/viya/open.html}{\url{https://www.sas.com/en_us/software/viya/open.html}}.
This presents opportunities for developers of smaller libraries with more QMC features, like BRODA and QMCPy, to connect their libraries with larger software environments. An example of QMCPy connecting with other software is given in the next subsection.

@@

\begin{figure}[t]
    \centering 
    \raisebox{0.5cm}{\includegraphics[width=0.4\textwidth]{BeamDrawing.png}}
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection (upper left, image taken from~\cite{ParSee22a} with permission \SCComment{to be obtained}) of a cantilevered beam  (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$ (lower right), and parallel processing gives a speed-up factor of four to five (lower left). These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{CDHJSfigbeam}
\end{figure}

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to an external load $f(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = f(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{f}$, where $\hat{u}$ and $\hat{f} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
QMCPy computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

\SCComment{Perhaps we could respond to referee's comment \#2 and \#4 here by framing it as future work.}

\section*{Acknowledgements}\SCComment{Added this section}
The authors are grateful to the referee's constructive comments.
